
---
title: "Exploring Data - Ch2"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
# __VERİ KEŞFİ__ (Exploring data)  
**This chapter covers**

- Using summary statistics to explore data

- Exploring data using visualization

- Finding problems and issues during data exploration

```{r}
install.packages("ggplot2")
library(ggplot2)
library(tidyverse)
library(party)
library(knitr)
library(hexbin)
library(naniar)
library(ROCR)
library(dplyr)
```


```{r}
custdata1 <- read.table("https://github.com/scizmeli/datascience_class_helper_files/raw/master/custdata.tsv",header=T,sep='\t') #control
```
```{r}
custdata <- read.table('https://github.com/scizmeli/datascience_class_helper_files/raw/master/custdata1.tsv',header=T,sep=',')
```
```{r}
head(custdata1) 
```
```{r}
knitr::kable(head(custdata1))
```
```{r}
summary(custdata1)
```
Income ranges from zero to over half a million dollars; a very wide range.!
```{r}
summary(custdata$income)


```
Spotting problems using graphics and visualization
-A graphic should display as much information as it can, with the lowest possible
cognitive strain to the viewer.

-Strive for clarity. Make the data stand out. Specific tips for increasing clarity
include

–Avoid too many superimposed elements, such as too many curves in the
same graphing space.
```{r}
plot(custdata$income, col = "red")
```
```{r}
hist(custdata$income, col = "green")
```
```{r}
library(ggplot2)
ggplot(custdata) +
geom_histogram(aes(x=age),
binwidth=5, fill="gray")
```
A histogram tells you where your data is concentrated. It also visually highlights
outliers and anomalies.
```{r}
#library(ggplot)
library(scales)
ggplot(custdata) + geom_density(aes(x=income)) +
scale_x_continuous(labels=dollar)
```
```{r}
ggplot(custdata)+
geom_density(aes(x=income))+
scale_x_log10(breaks=c(100,1000,10000,100000), labels=dollar)+
annotation_logticks(sides="bt")
```
```{r}
Income <- custdata$income/100
```
```{r}
ggplot(custdata) + geom_bar(aes(x=marital.stat), fill="lightblue")
# barplot kategorik degisken
```
```{r}
# barplot grafigi - eyaletler kategorik degisken
ggplot(custdata) + 
geom_bar(aes(x=state.of.res), 
fill="purple") +
coord_flip() +
theme(axis.text.y=element_text(size=rel(0.8)))
```
Producing a bar chart with sorted categories
```{r}
statesums <- table(custdata$state.of.res)
statef <- as.data.frame(statesums)
colnames(statef)<-c("state.of.res", "count")
summary(statef)
```
```{r}
statef <- transform(statef,state.of.res=reorder(state.of.res, count))
summary(statef)
```
```{r}
ggplot(statef)+ geom_bar(aes(x=state.of.res,y=count),
stat="identity",
fill="purple") +
coord_flip() +
theme(axis.text.y=element_text(size=rel(0.8)))
```
Producing a line plot
```{r}
x <- runif(100)
y <- x^2 + 0.2*x
ggplot(data.frame(x=x,y=y), aes(x=x,y=y)) + geom_line()
```
Examining the correlation between age and income
```{r}
custdata2 <- subset(custdata,(custdata$age > 0 & custdata$age < 100 & custdata$income > 0))
cor(custdata2$age, custdata2$income)
```
The negative correlation is surprising, since you’d expect that income should increase
as people get older. A visualization gives you more insight into what’s going on than a
single number can. Let’s try a scatter plot first; you plot figure
```{r}
ggplot(custdata2, aes(x=age, y=income)) +
geom_point() + ylim(0, 200000)
```
```{r}
ggplot(custdata2, aes(x=age, y=income)) + geom_point() + stat_smooth(method="lm") +
ylim(0, 200000)
# yasla gelir arasında dogrusal ilsiki yok cunku korelasyon degeri 0.02 cıktı.gayet çok az
```
In this case, the linear fit doesn’t really capture the shape of the data. You can better
capture the shape by instead plotting a smoothing curve through the data
```{r}
ggplot(custdata2, aes(x=age, y=income)) + geom_point() + geom_smooth() +
ylim(0, 200000)
# baska bir model kurarak dogrusal olmayan bir iliski vardır.
```
Plotting the distribution of health.ins as a function of age
```{r}
ggplot(custdata2, aes(x=age, y=as.numeric(health.ins))) +geom_point(position=position_jitter(w=0.05, h=0.05)) +
geom_smooth()
```
__HEXBIN PLOTS__
A hexbin plot is like a two-dimensional histogram. The data is divided into bins, and the
number of data points in each bin is represented by color or shading. Let’s go back to
the income versus age example.
```{r}
#install.packages("hexbin")
library(hexbin)
ggplot(custdata2, aes(x=age, y=income)) + geom_hex(binwidth=c(5, 10000)) +
geom_smooth(color="white", se=F) +
ylim(0,200000)
```
__BAR CHARTS FOR TWO CATEGORICAL VARIABLES__

Let’s examine the relationship between marital status and the probability of health
insurance coverage. The most straightforward way to visualize this is with a stacked bar
chart, as shown in figure 3.15.
```{r}
ggplot(custdata) + geom_bar(aes(x=marital.stat,
fill=health.ins))
```
```{r}
ggplot(custdata) + geom_bar(aes(x=marital.stat,
fill=health.ins),
position="dodge")
```
```{r}
ggplot(custdata) + geom_bar(aes(x=marital.stat,
fill=health.ins),
position="fill")
```
Plotting data with a rug
```{r}
ggplot(custdata, aes(x=marital.stat)) +
geom_bar(aes(fill=health.ins), position="fill") +
geom_point(aes(y=-0.05), size=0.75, alpha=0.3,
position=position_jitter(h=0.01))
```
Plotting a bar chart with and without facets
```{r}
ggplot(custdata2) +
geom_bar(aes(x=housing.type, fill=marital.stat ),
position="dodge") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
ggplot(custdata2) +
geom_bar(aes(x=marital.stat), position="dodge",
fill="darkgray") +
facet_wrap(~housing.type, scales="free_y") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
__EXTRA__
```{r}
library(naniar)
vis_miss(custdata)
#verileri eksik olan yerlerde çizgi cizilmis 
```
```{r}
gg_miss_fct(x = custdata, fct = sex)
```
```{r}
gg_miss_fct(x = custdata, fct = age)
```
```{r}
gg_miss_fct(x = custdata, fct = health.ins)
```
```{r}
gg_miss_fct(x = custdata, fct = num.vehicles)
```
